A Study of Residual Adapters for Multi-Domain Neural Machine Translation

### 2020-08-15 2020/wmt_20/emnlp2020.tex
rewrote section 3

\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.95}

@@

\section{Residual adapters \label{sec:res}}

@@

\subsubsection{Design space and variants}
This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.

@@

\subsection{Baseline architectures \label{ssec:baseline}}
\fyTodo{Write this - settings and parameters for Mixed and Full-FT}.
% Our baselines are standard for multi-domain systems.
% \footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.}
Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we train the following baselines:
\begin{itemize}
\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs. We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
\end{itemize}

@@

\subsection{Multi-domain systems}
we compare several proposals from the literature on multi-domain MT to finetuning method. The methods included in our experiments are described as follow
\begin{itemize}
\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
\item the three proposals of \newcite{Britz17mixing}. \system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}. The training uses reference tags and inference is performed with predicted tags, just like for regular target words. \system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations. These methods thus only use domain tags in training.
\item an original, multi-domain, version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where a domain-specific adaptation module is included in all the Transformer layers; within each layer, residual connections make it possible to by-pass this module. Contrarily to \cite{Bapna19simple}, we do not start with a trained generic system, but learn the multi-domain from scratch.\fyDone{Check this.}
\end{itemize}

Because of the limit of our computational resources, we limit our experiments in this section to one language pair En-Fr.

The table \ref{tab:performance-multi} shows that the full-finetuning outperforms all other systems in the in-domain tests. \system{FT-Block} is able to reduce the margin in several domains, showing the efficacy of residual adapters. Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}. However, \system{MDL-Res} is outperformed by \system{FT-Block}. It means that individually finetuning generic model is better than multi-task training. However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
   \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
   \hline 
    \system{DC-Tag}       & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}      & 37.7  & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9 \\
    \system{LDR}            & 37.0   & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8 \\
    \system{TTM}            & 37.3 & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7  \\
    \system{DM}             & 35.6 & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4 \\ 
    \system{ADM}           & 36.4 & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5\\
    \system{MDL-Res}     & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems in En-Fr. }
  \label{tab:performance-multi}
\end{table*}

\subsection{Varying the positions and number of residual adapters}
Tables \ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} display the performace of NMT model in 6 domains: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for the language pair En-Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En-De. The original version \system{FT-Block} outperforms the baseline \system{mixed} in almost all domains while stays equivalent to the baseline in \domain{med} in En-Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En-De. However, \system{FT-full} still stay as best model in all domain tests.

By varying the number and positions of residual adapters as aforementioned in section \ref{ssec:architecture}, we assess several variants of applying residual adapters. \fyTodo{Fix style here} Because the set of possible configurations is large, we only perform experiments on layers at position: 2, 4, 6 on each side of NMT model (encoder and decoder) which covers a wide range in the hierarchy of attention layers. To observe how much the number of residual adapters affects the performance of the adapted model. By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model. By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}

 In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{FT-Block$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{FT-Block$_{(2)}$}, \system{FT-Block$_{(4)}$} , \system{FT-Block$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyTodo{What does bold mean ?}\fyTodo{Number of parameters ?}

\begin{table*}
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 & 65M/0 \\
    \system{FT-Block}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\
    \system{FT-Block-WD}     & 37.18 & 55.99 & 52.93 & 33.36 & 46.03 & 90.65 & 52.7 & 65M/12M \\
   \system{FT-Block-LR}     & 37.45 & 56.09 & 51.84 & 33.29 & 45.02 & 89.7 & 52.2 & 65M/12M \\  
    \system{FT-Block$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
    \system{FT-Block$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
    \system{FT-Block$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
   \system{FT-Block$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\
     \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr). We report BLEU scores for each domain, as well as averages across languages. In column \domain{params}, we report number of domain-agnostic/domain-specific parameters} \fyDone{Boldface ?}
  \label{tab:performance-en-fr-pos-reg}
\end{table*}

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
    \system{Mixed}  & 29.31 & 30.48 & 17.64 & 38.11 & 47.94 & 20.95  & 30.59 & 213M/0M\\
   \system{FT-Block}     & 29.65 & 30.45 & 19.21 & 48.99 & 47.22 & 20.63 & 33.14 & 213M/96M \\
   \system{FT-Block-WD}     & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
   \system{FT-Block-LR}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M\\
   \system{FT-Block$_{(6)}$}     & 29.47 & 30.39 & 18.13 & 49.14 & 46.95 & 20.45 & 32.42 & 213M/16M \\
   \system{FT-Block$_{(4)}$}     & 29.69 & 30.4 & 18.07 & 49.61 & 47.05 & 20.64 & 32.58 & 213/16M\\
   \system{FT-Block$_{(2)}$}   & 29.64 & 30.4 & 18.29 & 49.41 & 46.71 & 20.59 & 32.51  & 213M/16M\\
   \system{FT-Block$_{(2,4,6)}$}  & 29.68  & 30.55 & 18.85 & 49.57 & 47.09 & 20.63 &  32.73 & 213M/48M \\
     \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains. In column \domain{params}, we report number of domain-agnostic/domain-specific parameters}
  \label{tab:performance-en-de-pos-reg}
\end{table*}

\fyTodo{Why HW worse than standard version ?} \fyTodo{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}

\subsection{Regularization of fine-tuning}
In the translation from English into German, we find that domains \domain{tour}, \domain{ecb} are extremely small and account for a very small percentage of the training data, with respective proportions of 0.039\% and 0.022\%. Fine-tuning on these domains can lead to serious overfitting. We assess two well-known regularization techniques for adapter modules, that could help avoid the overfitting issues: weight decay and layer regularization.

We choose optimal hyper-parameters: weight decay's coefficient, layer regularizer's coefficient $\lambda$ by grid search with a small set of values $\{ 10^{-3}, 10^{-4}, 10^{-5} \}$.

Results in Table~\ref{tab:performance-en-de-pos-reg} show that weight decay (\system{FT-Block-WD}) outperforms the more basic \system{FT-Block} while Layer regularization does not show any improvement compared to \system{FT-Block}.
\fyTodo{How is the weight decay parameter set ?}

@@

We assess the efficacy of proposed variants: Highway residual adapters \system{FT-HW-Block} and Gated residual adapter \system{FT-Gate-Block} by the same experiments in the previous section except that we only perform them in language pair En-Fr.

To measure the robustness against out-of-domain examples, we perform translations with wrong domain information to trick the systems that need to know domain a priori. In the table \ref{tab:performance-random}, the column \domain{RND} stores the score of multi-domain models, that use domain information to translate, in the situation in which we assign random domain tag to each test sentence. Besides the \system{RND} test, we use a new domain \domain{news} to assess the performance of systems in out-of-domain examples. Given a sentence of \domain{news}, we predict the closest domain to it by choosing the lowest perplexity of the sentence among those computed by in-domain language models and use the corresponding adapters to translate the sentence.

We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.

Table \ref{tab:performance-random} shows that variants \system{FT-HW-Block} and \system{FT-Gate-Block} achieve equivalent performances to the original version \system{FT-Block} of \cite{Bapna19simple}.

The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}. Our proposed model \system{FT-Gate-Block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter. \system{FT-Gate-Block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.

In the new domain test \domain{news}, generic model outperforms other systems. Finetuning systems including \system{FT-Full}, \system{FT-Block} and \system{FT-HW-Block} underperforms generic model by large margin. More regularized variants including \system{FT-Full-UEW}, and \system{FT-Gate-Block} are able to reduce the gap of performance compared to generic.

The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{FT-Block}.

>>>>

\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.8}

@@

\section{Residual adapters \label{sec:res}}
In this section, we describe the basic version of the residual adapter architectures, as well as two original variants of this model.

@@

\subsubsection{Design space and variants \label{sssec:design-space}}
This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.

@@

\subsection{Baseline architectures \label{ssec:baseline}}
\fyTodo{Write this - settings and parameters for Mixed and Full-FT}.
% Our baselines are standard for multi-domain systems.
% \footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.}
Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we train the following baselines:
\begin{itemize}
\item a generic model trained on a concatenation of all corpora, denoted \system{Mixed}\fyDone{Or mixed nat ?}
\item a fine-tuned model \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed} system, further trained on each domain with early stopping when the dev BLEU stops increasing during 3 consecutive epochs.
%  We again contrast two versions: full fine-tuning (\system{FT-Full}), which update all the parameters of the initial generic model \system{Mixed}; and the variants of \cite{Bapna19simple} (\system{FT-Block}).
\end{itemize}

@@

\subsection{Multi-domain systems}
In this section, we evaluate several proposals from the literature on multi-domain adaptation and compare them to full fine-tuning on the one hand, and to two variants of the residual adapter architecture on the other hand.

\begin{table*}[t!]
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Mixed}        & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  &  49.4 \\
    \system{FT-Full}       & 37.7 & 59.2 & 54.5 & 34.0 & 46.8 & 90.8 & 53.8 \\
    \hline 
    \system{DC-Tag}      & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5  & 50.1 \\
    \system{DC-Feat}     & 37.7 & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 & 49.9  \\
    \system{LDR}            & 37.0  & 54.7 & 49.9 & 33.9 & 43.6 & 79.9 & 49.8    \\
    \system{TTM}           & 37.3  & 54.9 & 49.5 & 32.9 & 43.6 & 79.9 & 49.7   \\
    \system{DM}            & 35.6  & 49.5  & 45.6 & 29.9 & 37.1 & 62.4 & 43.4   \\ 
    \system{ADM}          & 36.4  & 53.5  & 48.3 & 32.0 & 41.5 & 73.4 & 47.5   \\
    \hline
    \system{Res-Adap}         & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 \\ 
    \system{Res-Adap-MT}  & 37.9 & 56.0 & 51.2  & 33.5 & 44.4 & 88.3 & 51.9 \\
    \hline
  \end{tabular}
  \caption{Baseline translation performance of various multi-domain MT systems (En$\rightarrow$Fr).}
  \label{tab:performance-multi}
\end{table*}

The reference methods included in our experiments are the following:
\begin{itemize}
\item a system using ``domain control'' \cite{Kobus17domaincontrol}. In this approach, domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}) or in the form of a supplementary feature for each word (\system{DC-Feat});
\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain-specific part (\system{LDR});
\item the three proposals of \newcite{Britz17mixing}. \system{TTM} is a feature-based approach where the domain tag is introduced as an extra word \textsl{on the target side}. The training uses reference tags and inference is performed with predicted tags, just like for regular target words. \system{DM} is a multi-task learner where a domain classifier is trained on top of the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations. These methods thus only use domain labels in training.
\end{itemize}

The two variants of the residual adapter model included in this first round of experiment have presented in Section~\ref{ssec:architecture}: \system{Res-Adap} is the multi-domain version of the approach of \citet{Bapna19simple} based on a two step training procedure; while \system{Res-Adap-MT} is the ``multi-task'' version, where the parameters of the generic model and of the adapter  are learned simultaneously from scratch.

Because of the limit of our computational resources, we restrict our experiments in this section to the En$\rightarrow$Fr task. Results are in Table~\ref{tab:performance-multi}.

These results first show that full fine-tuning outperforms all other methods for the in-domain test sets. However, \system{Res-Adap} is able to reduce the gap with this approach for several domains, showing the effectiveness of residual adapters. The ``multi-task'' variant is slightly less effective in our experiments than the basic version, where optimization is performed in two steps. As it turns out, using residual adapters prove here to be much more effective on average than the other reference multi-domain systems; it is also much better than the generic system for translating data from known domains, outperforming the \system{Mixed} system by more than 4 BLEU points in average. Gains are especially large for small domains such as \domain{law} and \domain{rel}.

% Multi-task training system \system{MDL-Res} gains important improvement compared to baseline \system{Mixed}. However, \system{MDL-Res} is outperformed by \system{Res-Adap}. It means that independently fine tuning the generic models is better than multi-task training. However, \system{MDL-Res} is better than other methods in multi-domain learning including:  \system{DC-Tag},  \system{DC-Feat}, \system{LDR}, \system{TTM}, \system{DM}, \system{ADM}.

\subsection{Varying the positions and number of residual adapters}

Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} report the BLEU scores for 6 domains in each language pair: \domain{med},\domain{law},\domain{bank},\domain{talk},\domain{it} and \domain{rel} for En$\rightarrow$Fr; \domain{gov}, \domain{eco}, \domain{tour}, \domain{bank}, \domain{med} and \domain{news} for En$\rightarrow$De. We first see that for the latter direction, the basic version \system{Res-Adap} also outperforms the \system{mixed} baseline on average, with large gains for the small domains \domain{tour}, \domain{bank} and comparable results for the other domains.
% in almost all domains while stays equivalent to the baseline in \domain{med} in En$\rightarrow$Fr and \domain{news}, \domain{gov}, \domain{eco}, \domain{med} in En$\rightarrow$De.
   %    However, \system{FT-full} still remains the best model for all test domains.
\fyFuture{Full FT for en->DE ? }

By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementation of residual adapters. \fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder). Two settings are considered: keeping just one adapter, or keeping the three. The trend is the same for the two language directions: suppressing adapters always decrease the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one. With only one adapter active, we observe small, unsignificant changes in performance. Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.

% which covers a wide range in the hierarchy of attention layers. To observe how much the number of residual adapters affects the performance of the adapted model. By the limit of computation resources, we limit our study to the case of 3 adapters at positions: 2,4,6 in each side of the NMT model. By doing this, we could cover 3 situations: 1 adapter (mentioned in the previous paragraph), 3 adapters and 6 adapters (in the normal setting) in each side (encoder and decoder).\fyTodo{We have 12 layers, have we not ?}

%  In most cases, the performance increases with respect to the number of residual adapters used in architecture. The setting \system{FT-block} using residual adapters at all levels outperforms setting \system{Res-Adap$_{(2,4,6)}$} using residual adapter at 3 levels, which outperforms \system{Res-Adap$_{(2)}$}, \system{Res-Adap$_{(4)}$} , \system{Res-Adap$_{(6)}$} using residual adapter at only one level. However, the difference in performance between positions is not significant except the case of domain \domain{rel} (En-Fr) in which the lower position shows the lower performance.\fyDone{What does bold mean ?}\fyDone{Number of parameters ?}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline 
    \system{Mixed}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 49.4 & 65M/0 \\
    \system{Res-Adap}     & 37.3 & 57.9 & 53.9 & 33.8 & 46.7 & 90.2 & 53.3 & 65M/12M\\ \hline
    \system{Res-Adap$_{(2,4,6)}$}     & 37.7 & 57 & 53 & 33.3 & 45 & 90 & 52.7 & 65M/6M\\
    \system{Res-Adap$_{(6)}$}     & 37.7 & 55.8 & 51.5 & 33.9 & 43.6 & 89.2 & 51.9 & 65M/2M \\
    \system{Res-Adap$_{(4)}$}     & 37.9 & 55.6 & 51.7 & 33.7 & 44.4 & 88.7 & 52 & 65M/2M\\
    \system{Res-Adap$_{(2)}$}     & 37.8 & 55.5 & 51.4 & 34 & 43.8 & 86.7 & 51.5 & 65M/2M\\ \hline
    \system{Res-Adap-WD}     & 37.2 & 56.0 & 52.9 & 33.4 & 46.0 & 90.6 & 52.7 & 65M/12M \\
    \system{Res-Adap-LR}      & 37.4 & 56.1 & 51.8 & 33.3 & 45.0 & 89.7 & 52.2 & 65M/12M \\  
     \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$Fr). We report BLEU scores for each domain, as well as averages across languages. Column \domain{params} reports the number of domain-agnostic/domain-specific parameters \label{tab:performance-en-fr-pos-reg}} \fyDone{Boldface ?}
\end{table*}

\begin{table*}[htbp]
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{gov}} & \multicolumn{1}{c|}{\domain{eco}} & \multicolumn{1}{c|}{\domain{tour}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{ med }} & \multicolumn{1}{c|}{\domain{ news}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c|}{\domain{params}} \\ \hline
    \system{Mixed}          & 29.3 & 30.5 & 17.6 & 38.1 & 47.9 & 20.9  & 30.6 & 213M/0M\\
   \system{Res-Adap}     & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1 & 213M/96M \\ \hline
    \system{Res-Adap$_{(2,4,6)}$} & 29.7  & 30.5 & 18.8 & 49.6 & 47.1 & 20.6 &  32.7 & 213M/48M \\ \hline
    \system{Res-Adap$_{(6)}$}      & 29.5 & 30.4 & 18.1 & 49.1 & 46.9 & 20.4 & 32.4 & 213M/16M \\
   \system{Res-Adap$_{(4)}$}       & 29.7 & 30.4 & 18.1 & 49.6 & 47.0 & 20.6 & 32.6 & 213/16M\\
   \system{Res-Adap$_{(2)}$}       & 29.6 & 30.4 & 18.3 & 49.4 & 46.7 & 20.6 & 32.5  & 213M/16M\\
    \system{Res-Adap-WD}         & 29.7 & 30.8 & 20.4 & 50.2 & 47.7 & 20.6 & 33.2 & 213M/96M \\
    \system{Res-Adap-LR}           & 29.6 & 30.4 & 19.2 & 49.0 & 47.2 & 20.6 & 33.1  & 213M/96M\\
    \hline
  \end{tabular}
  \caption{Translation performance of various fine-tuned systems (En$\rightarrow$De). We report BLEU scores for each domain, as well as averages across domains. In column \domain{params}, we report the number of domain-agnostic/domain-specific parameters}
  \label{tab:performance-en-de-pos-reg}
\end{table*}

\subsection{Regularizing fine-tuning \label{ssec:regularization-exp}}

The translation from English into German includes two domains (\domain{tour} and \domain{bank}) that are extremely small and account only for a very small fraction of the training data (respectively for 0.039\% and 0.022\% of the total number of sentences). Fine-tuning on these domains can lead to serious overfitting. We assess two well-known regularization techniques for adapter modules, that could help mitigate this problem: weight decay and layer regularization.

For each method, the optimal hyper-parameters (weight decay or layer regularization coefficient, see Section~\ref{sssec:design-space}) $\lambda$ by chosen by grid search among a small set of values ($\{ 10^{-3}, 10^{-4}, 10^{-5} \}$).

Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively affect the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains. Using layer regularization proves here to be comparatively less effective. Finding better ways to set the regularization parameters, for instance by varying $\lambda$ for each domain based on the available supervision data, is left for future work.  
\fyDone{How is the weight decay parameter set ?}
\fyDone{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}

@@

We now turn to the evaluation of our new architectural variants: Highway residual adapters \system{Res-Adap-HW} on the one hand, and Gated residual adapters \system{Res-Adap-Gated} on the other hand. We use the same domains and settings as before, focusing here exclusively on the language direction En$\rightarrow$Fr.

To also evaluate the robustness with respect to out-of-domain examples, we perform two additional experiments. We first generate translations with erroneous (more precisely: randomly assigned) domain information: the corresponding results appear in Table~\ref{tab:performance-random} under the column \domain{rnd}. We also compute translation for a domain unseen in training (\domain{news}) as follows. For each sentence of this test set, we automatically evaluate the closest domain\footnote{As measured by the perplexity of a language model trained with only in-domain data.\fyFuture{More on this for reproducibility}.}, then use the predicted domain label to compute the translation. This is an error prone process, which also challenges the robustness of our multi-domain systems. Results are in Table~\ref{tab:performance-random}.
\fyFuture{This is a baseline, and we have not discussed it - keep for next time: We also report model \system{FT-Full-UEW} which is obtained by full-finetuning the generic model using uniform elastic weight to avoid catastrophic forgetting.}

A first observation is that for domains seen in training, our variants \system{Res-Adap-HW} and \system{Res-Adap-Gated} achieve BLEU scores that are on a par to those of the original version (\system{Res-Adap}), with insignificant variations across test sets. \fyFuture{Comment on processing time}

The two other settings are instructive in several ways: they first clearly illustrate that the brittleness of domain-adapted systems, for which large drops in performance (more than 15 BLEU points on average) are observed when the domain label is randomly chosen. Our gated variant however proves much more robust than the other adaptation strategy, and performs almost on par to the generic system for that test condition. The same trend holds for the unseen \domain{news} domain, with \system{Res-Adap-Gated} being the best domain adapted system in our set, outperforming the other variants by about 2 BLEU points.

% The phenomenon of Catastrophic Forgetting manifests clearly in this situation where the performance of all multi-domain systems drops dramatically by approximately 10 points BLEU compared to generic model \system{Mixed}. Our proposed model \system{FT-Gate-Block} maintains equivalent performance to the generic model in \system{RND} thanks to the application of adaptive weight on the output of the residual adapter. \system{FT-Gate-Block} avoids relying on the residual adapters when predicting examples that come from domain far from the domain of the adapters.

% In the new domain test \domain{news}, generic model outperforms other systems. Finetuning systems including \system{FT-Full}, \system{Res-Adap} and \system{FT-HW-Block} underperforms generic model by large margin. More regularized variants including \system{FT-Full-UEW}, and \system{FT-Gate-Block} are able to reduce the gap of performance compared to generic.

% The column \domain{PROC} shows that highway residual adapter \system{FT-HW-Block} process examples faster than the original version \system{Res-Adap}.

% \fyTodo{Why HW worse than standard version ?}
